---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r read data}
library(utils)
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")
```

```{r model 1}
#first letter to a numeric alphabetic position.
data1 <- data[(data$dateRep == "11/08/2020"), ]
CountryAb <- as.integer(as.factor(substr(data1$countriesAndTerritories, 1, 1)))

#build a linear model for the relationship between Cumulative_number_for_14_days_of_COVID.19_cases_
model1 <- lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ CountryAb, data = data1)
qplot(CountryAb, Cumulative_number_for_14_days_of_COVID.19_cases_per_100000, data = data1)
summary(model1)
```

```{r model 2}
data2 <- data[(data$dateRep == "11/09/2020"), ]
CountryAb <- as.integer(as.factor(substr(data2$countriesAndTerritories, 2, 2)))

model2 <- lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ CountryAb, data = data2)
summary(model2)

```
```{r model 3}
library(tidyverse)
library(ggplot2)
data3 <- data
data3$weekday_no <- as.numeric(strftime(as.Date(data3$dateRep, "%d/%m/%Y"), "%u"))

data3_Thu <- data3 %>%
  filter(weekday_no == 1, continentExp == "Asia")

#CountryAb <- as.integer(as.factor(substr(data3$countriesAndTerritories, 2, 2)))

model3 <- lm(deaths ~ day, data = data3_Thu)
#qplot(day, cases, data = data3_Thu)
summary(model3)

```

```{r model 4}

data4 <- data[(data$dateRep == "05/05/2020"), ]
CountryCode <- as.integer(as.factor(substr(data4$countryterritoryCode, 2, 2)))

model4 <- lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ CountryCode, data = data4)
qplot(CountryCode, Cumulative_number_for_14_days_of_COVID.19_cases_per_100000, data = data4)
summary(model4)

```

```{r model 5}

data5 <- data
data5$weekday_no <- as.numeric(strftime(as.Date(data5$dateRep, "%d/%m/%Y"), "%u"))

data5_Poland <- data5 %>%
  filter(countriesAndTerritories == "Poland")

model5 <- lm(deaths ~ weekday_no, data = data5_Poland)
summary(model5)

```

** There is a positive significant correlation (p = 0.0408) between the day of a week and number of deaths caused by COVID-19 in Poland. This correlation doesn't seem to occur for many of the other european countries, with focus on slavic countries and countries with a similar population - it seem to be accidential. In other words, in order to present this correlation, the country has been cherry-picked from the pool of other countries, that don't show any similar correlation. **
